Regularization Techniques for PSF--Matching Kernels. II. Pre--Filtering Input Images

### 2014-12-20 submit/ms.tex
First take a modifying internal document for ApJ format.  About halfway though the first pass

\documentclass[prd, nofootinbib, floatfix, 11pt,tightenlines,times]{article}

% latex W13report.tex; bibtex W13report; latex W13report; latex W13report; dvips W13report; ps2pdf W13report.ps ; gv W13report.pdf

\usepackage[paperwidth=8.5in,paperheight=11in,centering,margin=1in]{geometry}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amsbsy}
\usepackage{natbib}
\usepackage{rotating}
\input epsf
\usepackage{amsmath}
\usepackage{caption}
\usepackage{wasysym}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{color}
%\usepackage{ulem}
%\usepackage{epstopdf}
\usepackage[toc,page]{appendix}

%\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\makeatletter
\newcommand*{\myfnsymbolsingle}[1]{%
  \ensuremath{%
    \ifcase#1% 0
    \or % 1
      \mathparagraph%   
    \or % 2
      \dagger
    \or % 3  
      \ddagger
    \or % 4   
      \mathsection
    \or % 5
      \mathparagraph
    \else % >= 6
      \@ctrerr  
    \fi
  }%   
}   
\makeatother
\newcommand*{\myfnsymbol}[1]{%
  \myfnsymbolsingle{\value{#1}}%
}
\renewcommand*{\thefootnote}{%
  \myfnsymbolsingle{\value{footnote}}%
}

%\renewcommand{\baselinestretch}{0.98}
\usepackage{epsfig}
\usepackage{titletoc}

% HOW TO SET UP AN 8.5 x 11:
% http://www.pages.drexel.edu/~pyo22/students/latexRelated/latexTutorial.html
\topmargin -1.5cm        % read Lamport p.163
\oddsidemargin -0.04cm   % read Lamport p.163
\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
\textwidth 16.59cm
\textheight 21.94cm 
\parskip 7.2pt           % sets spacing between paragraphs
\parindent 0pt       % sets leading space for paragraphs

@@

\author{
Andrew Becker (UW), Simon Krughoff (UW), Andrew Connolly (UW), \\
Russell Owen (UW), Paul Price (Princeton), Robert Lupton (Princeton), \\
Dustin Lang (CMU), Jim Bosch (Princeton), Robyn Allsman (LSSTC), \\
Yusra Al Sayyad (UW), Jacek Becla (SLAC), Steve Bickerton (Kavli IPMU), \\
Scott Daniel (UW), Greg Daues (NCSA), Gregory Dubois-Felsmann (SLAC), \\
Mike Freemon (NCSA),Perry Gee (UC Davis), Bill Glick (NCSA),\\
Steve Groom (IPAC), Mario Juric (LSSTC),Jeff Kantor (LSSTC),\\
Ron Lambert (NOAO), Kian-Tat Lim (SLAC),Bruce Mather (NCSA), \\
Serge Monkewitz (IPAC),Knut Olsen (NOAO), Steve Pietrowicz (NCSA), \\
Ray Plante (NCSA),Dick Shaw (NOAO), Douglas Smith (SLAC), \\
Schuyler Van Dyk (IPAC), and Daniel Wang (SLAC)\\ 
{\it for LSST Data Management}
\vspace{-2ex}}
\title{Report on Late Winter2013 Production: Image Differencing\vspace{-2ex}}
\date{\today}

\begin{document}

\maketitle
\vspace{-2ex}
We provide here a technical report on the LSST Data Management
late--Winter 2013 production runs.  This production focused on the
evaluation of LSST DM image differencing code and, in particular, 
the rate of false detections the pipeline generates.  As
inputs, we used a suite of simulated images generated in semi-idealized
(stars only, no saturation, cosmic rays, or defects) conditions and containing no intrinsic variability.  
Thus {\it any}
sources detected in these image differences are ``false positives''.

We investigated two distinct operational modes of the pipeline.  The
first pre--filters the science image with its own PSF, yielding a
point--source maximum likelihood image.  The template is then
PSF--matched to this image, subtracted from it, and detection and
measurement happen directly on this difference.  The second mode
matches the template image to the science image, creates their
difference, post--filters the difference with the science image PSF,
and then runs detection on this maximum likelihood image.  Measurement
happens on the detected footprints in the difference image.

We investigated the complexity of the basis functions used to model
the PSF--matching kernel, and the degree to which this kernel varies
across the images.  We generate several goodness of fit statistics for
a control sample of sources that were {\it not} used in the kernel fit
(notably the mean squared error), in order to assess the interpolation ability
of the matching kernel.  These metrics are shown to correlate with the
numbers of false positives.  We calculate a mean squared error metric
that shows a minimum value close to the configurations that yielded
the lowest number of false positives, and shows promise as a means to
algorithmically choose the optimal pipeline configuration.

We compared the measured numbers of false detections to the number we
expect due to random fluctuations in a background Gaussian field.  The
latter is a function of both seeing in the images, and of the
detection threshold used.  At the fiducial detection threshold of
5--sigma, the number of expected false positives per LSST sensor is
3--12, depending upon seeing (see Section \ref{sec-analyticfp}).  
We measure a similar number in the
difference images, indicating that we are able to achieve the
systematic floor in false detection rate.  We examined our measured
false detection rate as a function of detection threshold and find
that it scales as expected.  From these measures, we estimate that in
order for there to be less than 1 expected false detection per sensor
(i.e.\ $\sim$20\% of the alert production rate) we must operate at a
detection threshold of 5.5--sigma (for seeings as good as 0.6''). This
result is seeing dependent.

We found a significant dependence of the false positive rate on our
understanding of the variance; even a 2\% misestimate of the noise may
change the false positive rate by a factor of 1.5--2 at 5--sigma.
Pre--filtering is preferred in this regard, as our variance planes are
shown to track the variance to higher fidelity at the detection stage.
We find that over/under subtraction of the image background leads to a
bias in the positive to negative false detection ratio, but not on the
overall number of detections.  Pre--filtering may obviate the need for
deconvolution of the template image when the science image has better
seeing, which is a clear advantage of the technique.

We outline several additional steps we will pursue on the Summer 2013
timescale, including operating on images with more astrophysical and
atmospheric complexity, improving measurement on PSF--filtered
sources, adopting Gaussian processes for modeling of the spatial
kernel, and using statistics calculated during the image subtraction
to tune the algorithm ``on the fly'' (i.e. predict the numbers of
false positives it will generate).

Please address any questions about this document to dm-help@lsstcorp.org.

\clearpage
\tableofcontents
\clearpage

\section{Production Scope and Goals}

The focus of this production was to optimize and quantify the false
detection rate in the PSF--matched subtraction of a science image and
a reference image (referred to as the template image below).  We used
simulated images generated using semi-ideal conditions, and with no
intrinsic variability in the sources.  The metric we optimized is the
number of false detections that are detected and measured on the image
difference.  As the input were simulated with no variability, any
detections found in the difference images are ``false positives''.  We
calculated several additional quality metrics during the image
subtraction process, and correlated them with the number of false
detections, as a means to predict the number of false detections we
may expect downstream from the image subtraction.  Once we found an
optimum configuration, we perturbed the solution along multiple
dimensions to understand how the false detection rate responded.

\subsection{Algorithm and Background}

We outline the image subtraction process below.  We assume that science
image $S(x,y)$ can be modeled as a convolution of the template image
$T(x,y)$ by a PSF--matching kernel $K(u,v;x,y)$ (indices $u,v$ indicate
that the kernel itself is a 2--dimensional function, which varies as a
function of position $x,y$ in the image.  During convolution and correlation
there is an implicit summation over $u,v$).  The images will have
different point spread functions (PSFs), which are the time--averaged
transfer functions of a point source through the Earth's atmosphere,
telescope optics, and into the silicon of the detector before being
read out.  The essence of image subtraction is to match the PSFs of
these two images so that they may be subtracted pixel by pixel.  We
further assume that the PSF--matching kernel may be decomposed using a
set of basis functions $K(u,v) = \sum_i a_i K_i(u,v)$, where the
coefficients in front of each basis are determined through:
%
\begin{eqnarray}
C_i & \equiv & (K_i \otimes T) \\ \nonumber
b_{i}  & = & \sum_{x,y} {{C_i(x,y) S(x,y)}\over{\sigma^2(x,y)}}   \nonumber \\ 
M_{ij} & = & \sum_{x,y} {{C_i(x,y) C_j(x,y)}\over{\sigma^2(x,y)}}  \nonumber \\ 
a_{i}  & = & M^{-1}_{ij} b_{j} \nonumber 
\label{eq-soln}
\end{eqnarray}
where $\sigma^2(x,y)$ is the per--pixel variance stored in the {\tt
  variance} plane of each LSST {\tt exposure}.  To generate a
spatially varying model for the kernel, we assume that the relative
weights of the basis functions $a_i$ themselves vary spatially,
i.e. $K(u,v;x,y) = \sum_i a_i(x,y) K_i(u,v)$.  We also assume that
there is a differential background between the two images $b(x,y)$
that may be fit for using a low--order Chebyshev polynomial.  The
image difference is then calculated through $D(x,y) = S(x,y) - T(x,y)
\otimes K(u,v;x,y) - b(x,y)$.

The basis functions themselves $K_i(u,v)$ are a degree of freedom in
this problem.  The original implementations \citep{Alard98,Alard00}
used a set of 3 Gaussians (represented by our {\tt Config} variable
nGauss), each with a different width (sigGauss), and each modified by
a Laguerre polynomial to a given order (degGauss).  Subsequent studies
\citep[e.g.][]{2007AN....328...16I} have suggested that a constant
ratio be maintained between the different Gaussian widths, such that
$\sigma_{i+1} = \beta \times \sigma_{i}$.  We use the value $\beta =
2.0$ for this production.  We set the overall scale for sigma by
noting that, under the assumption that the PSFs of the images are
Gaussian ($\sigma_S$ for the science image and $\sigma_T$ for the
template image), the sigma of the matching kernel should be simply
$\sigma_K^2 = \sigma_S^2 - \sigma_T^2$.  We use this width for the
central Gaussian in the basis sequences where we use more than one
Gaussian, or the sole Gaussian in the sequences when we use only one.

Detection on the difference image occurs through correlation of
$D(x,y)$ with the science image's PSF, yielding optimally filtered
detection image $D'(x,y) = D(x,y) \circ PSF_S(u,v;x,y)$ where $\circ$
denotes correlation (currently the DM stack uses convolution instead of correlation,
reported as ticket \#2552).  The values of the pixels in $D'(x,y)$ provide a
maximum likelihood estimate of there being a point source at that
position.  Detection occurs by simply finding pixels that are more
than N sigma above the square root of the per--pixel variance.

In this production, we investigated two orders of operations for the
optimal filtering by the PSF.  In the first method, called
pre--filtering, we PSF--match the template image to a pre--filtered
science image:
\begin{eqnarray}
D_{Pre}(x,y) & = & S(x,y) \circ PSF_S(u,v;x,y) - T(x,y) \otimes K(u,v;x,y) \nonumber \\ 
D'_{Pre}(x,y) & = & D_{Pre}(x,y). \nonumber 
\end{eqnarray}
By pre--smoothing the image that we are matching the template to --
$S(x,y) \circ PSF_S(u,v;x,y)$ -- we more frequently avoid the case of
PSF--matching {\it deconvolution}, where the full width at half maximum (FWHM)
 of $T$ is broader
than $S$.  This pre--filtering increases $\sigma_S$ by a factor of $\sqrt{2}$.
The PSF matching kernels will need to be correspondingly larger to
account for the larger effective PSFs.  We are able to run detection
directly on this image with different measurement algorithms (see Section \ref{sec-meas}).

In the second, which is the classical post--filtering implementation of
the technique, we create a difference image that is then
post--filtered with the PSF for detection:
\begin{eqnarray}
D_{Post}(x,y) & = & S(x,y) - T(x,y) \otimes K(u,v;x,y) \nonumber \\ 
D'_{Post}(x,y) & = & D_{Post}(x,y) \circ PSF_S(u,v;x,y).  \nonumber 
\end{eqnarray}

%One complication is that measurement on $D'_{Pre}$ is necessarily
%different than on $D'_{Post}$.

@@

Specialized simulations were developed with constrained astrophysical
complexity for the initial test data. The purpose of these images was
to define a baseline for how well image subtraction would perform in
the absence of chromatic effects (e.g.\ differential chromatic
refractions), and perturbations in the optical system. To accomplish
this the data comprised $i$-band simulations of a random distribution
of stars with a density of $\sim 1000$ stars per CCD and with an
equivalent $r$-band magnitude range ($19<r<21$). All stars in this
sample had identical spectral energy distributions
(km50\_5000.fits\_g20\_5140.gz; see the phosim page,{\tt http://ls.st/1vo} for links to the
spectra), and, for all sources, variability,
proper motion, and parallax were set to zero. All simulations were
made at the zenith.

Nine sensors of the central raft were simulated, for a total of 27
difference images per configuration to use in our comparative
analysis.  In each case a pair of snap exposures was generated, however only
the first snap was used in the analysis.  Each simulation run was generated using the following
configurations:
\begin{itemize}
\item Visit 6866600: A pair of 300s exposure template images with
  an atmospheric seeing value of 0.88 arcsec.
\item Visit 6866601: A pair of 15s exposure exposures with 0.6
 arcsec atmospheric seeing.
\item Visit 6866602: A pair of 15s exposure exposures with 0.88
 arcsec atmospheric seeing.
\item Visit 6866603: A pair of 15s exposure exposures with 1.2
 arcsec atmospheric seeing.
\end{itemize}

For each configuration, v3.2.5 of the LSST phosim package (see {\tt http://ls.st/1vo} for usage)
 was used to
generate the data. The sky background was dark and set to produce
$\sim$150 counts per pixel, with airglow variations set to
zero. Telescope and camera perturbations, tracking errors, field
rotation, and saturation and blooming effects were all disabled. To
enable the use of eimages directly within the analysis (i.e.\ to
remove the need for the construction of a full CCD from the amplifier
images, calibration of the simulated data, and the generation of
photometric zero points) the pixel quantum efficiency variations were
set to zero, cloud screens were turned off and no cosmic rays were
generated.  The phosim commands file for the generation of the
simulated data is listed in Appendix~\ref{sec-phosim}.  Reference
catalogs and astrometry.net indices were constructed from the input
catalogs.  A secondary set of simulation images was generated for an
extended Summer 2013 analysis.  These data are described in
Appendix~\ref{sec-phosimextend}.

\section{Work Performed}

Major tasks that were completed in the run--up to production include:
implementing pre--filtering for image subtraction (ticket \#2551);
implementing source selection for image subtraction (\#2554, \#2596);
implementing measurement for difference image source ({\tt DiaSource})
characterization (\#2595); implementing RegisterTask for
image--to--image registration (\#2562, \#2650); defining and
implementing metrics for testing ip\_diffim (\#2636); designing a tool
to scrape metadata from data repositories (\#2640); and implementing
flux measurement for PSF--filtered sources (\#2754).

Below we describe in detail the overall pipeline that was established
as the results of these tasks.  We follow with in--depth analysis of
implemented algorithms that are central to this production.

@@

At the beginning of the pipeline, we have as an input the science
image $S(x,y)$ as a {\tt calexp} instance.  The bounding box of the
science image, in celestial coordinates, is ideally used to query the
coadd repository to return the coadd patches that overlap the image.
However, because of difficulties in fitting absolute astrometry to the
images (Section~\ref{subsec-astrom}), we instead took the approach of
directly using the {\tt calexp} of the deep visit (v6866600) as the
reference visit (as opposed to taking this deep {\tt calexp} and
turning it into a coadd).

Next, we used the DiaCatalogSourceSelector to query the ImSim
reference catalog for appropriate sources to use for PSF matching.
This selector allows the user to specify the brightness and color
range of the objects, toggle star or galaxy selection, and to include
variable objects or not.  In this semi-ideal case we use all sources.
We selected a fraction (1/5) of these
objects to serve as a control sample, in order to assess the
effectiveness of $K(u,v;x,y)$ to subtract objects that were {\it not}
used in the least squares fit.

The reference and science {\tt calexps} were then astrometrically
registered to each other, with the resampling operations happening on
the reference visit as it has the larger signal--to--noise (S/N).
Object--to--object matching, done using the source list above
cross--matched with a source list extracted from the coadd, was used
to provide a relative astrometric transformation between the two
images.  RegisterTask was used to do the resampling.  The resulting
warped reference image was what we use as the template image $T(x,y)$.
The residuals of the object--to--object match were persisted as part
of the debugging infrastructure.  We assumed that the background
subtraction on the input {\tt calexps} removed the high order features
in the backgrounds; therefore we simultaneously fit for a 1st--order
spatially varying differential background $b(x,y)$ along with the
kernel coefficients.

Images $S(x,y)$ (optionally pre-filtered with $PSF_S(u,v;x,y)$) and
$T(x,y)$ were sent to ImagePsfMatchTask, whose purpose is to fit for
$K(u,v;x,y)$ and to produce the difference image $D(x,y)$.

%\subsubsection{Pipeline optimization}
\begin{figure}[!hpb]
\centering{\includegraphics[width=0.8\textwidth]{figures/basis3.eps}}
\caption{Example basis set with nGauss=3 (highlighted in red),
  sigGauss=(0.5 sig, 1.0 sig, 2.0 sig), and degGauss=4, yielding
  Laguerre polynomials of order (4,2,2).}
\label{basis}
\end{figure}

 The first step is to use the
PSF Gaussian widths to define the {\it sizes} of the Gaussians' shapes
in $K_i(u,v)$.  As mentioned above, the main width was chosen to be
$\sqrt{\sigma_S^2 - \sigma_T^2}$.  We investigated variations on the
configuration where we used one Gaussian with the above width, and
where we used three Gaussians, the central of which has this width and
the others being a factor of $\beta = 2.0$ smaller and larger.  We
modified each of the Gaussians by a set of Laguerre polynomials of a
given order.  The smallest Gaussian was modified to the specified
order, with the others modified by floor(order/2).  Figure~\ref{basis}
provides an example basis set with this configuration.  The total
number of shapes in each basis were: $\sum_i^{\rm nGauss} ({\rm
  degGauss}_i+1)\times({\rm degGauss}_i+2)/2$ \footnote{ A notable
  exception to this algorithm is found in the case where $\sigma_T >
  \sigma_S$, i.e. a sharpening must be done to the template to match
  it to the science image.  This is the case of deconvolution.  A
  theoretical prescription for basis design in this case is given in
  \cite{0266-5611-26-8-085002}.  In our implementation of this algorithm, nGauss is {\it
    always} fixed to value 3, and degGauss is always fixed to 3.  The
  widths of the Gaussians are determined through the algorithm
  specified in \cite{0266-5611-26-8-085002}, using as inputs the
  sequence of Gaussians that we would have used to match a Gaussian of
  width $\sigma_S$ to $\sigma_T$ (i.e. as if we would have convolved
  the science image and not the template image).  There are a variety
  of reasons that actually operating on the science image is
  undesirable, including: the science image is lower S/N than the
  template; the science image is likely to have masked defects,
  including cosmic rays and bad pixels, whose reach would be expanded
  due to the convolution; unmasked features such as faint cosmic rays
  are more likely to look PSF--like after convolution (cosmic rays
  will likely have been sigma clipped out of the coadd).  However, it
  is not clear that deconvolving the template image is in principle
  any better.  A far better option, which we explore here, is to
  pre--filter the science image, which will in the vast majority of
  cases provide a larger PSF (by a factor of $\sqrt{2}$) to match the
  template to.  This is the case for v6866601 in this study, where the
  post--filtered analysis yields deconvolution of the template, while
  the pre--filtered analysis does not.  }.
The dimensions of the PSF matching kernel were chosen to be 6 times
the largest Gaussian width, with a minimum kernelSize of 21x21 pixels.
For kernels that are significantly smaller than this, the Gaussians
have significant (non--zero) power at the kernel boundaries, leading
to square systematic artifacts at the scale of the kernel in the
difference images (see Figure~\ref{fig_galleryBad}).
\begin{figure}[!hpb]
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/deconv2.eps} 
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/size2.eps} \\
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/order2.eps} 
\includegraphics[width=0.5\textwidth, height=0.25\textwidth]{figures/shape2.eps} \\
\caption{Four example panels showing failure modes of the image
  subtraction software.  Each panel itself shows a 2x2 pair of
  {\tt KernelCandidate} difference images; on the {\it left} subpanel is the residual from

>>>>

\documentclass[floatfix, apj]{emulateapj}

% latex ms.tex; bibtex ms; latex ms; latex ms; dvips ms; ps2pdf ms.ps ; gv ms.pdf
\usepackage[paperwidth=8.5in, paperheight=11in, centering, margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{epsfig}

@@

\begin{document}
\title{Regularization Techniques for PSF--Matching Kernels. II. Pre--Filtering Input Images}

\author{
A.C.~Becker\altaffilmark{1},
R.H.~Lupton\altaffilmark{2},
K.S.~Krughoff\altaffilmark{1},
et al.
}
\altaffiltext{1}{Astronomy Department, University of Washington, Seattle, WA 98195}
\altaffiltext{2}{Department of Astrophysical Sciences, Princeton University, Princeton, NJ 08544}

\date{\today}

\begin{abstract}

We report on a novel technique for reducing the numbers of false detections in difference images due to deconvolution (or sharpening) by the profile--matching kernel.
In this technique, we pre--smooth (or pre--filter) the input science image by its own point spread function before performing profile--matching.
This results in a maximum--likelihood image for point source detection, which has a $\sqrt{2}$ broader point--source profile than the image PSF.
We then match a reference image to this smoothed image, effectively resulting in a point--source maximum likelihood difference image.
The process of detection happens directly on this image; however, measurement techniques must be modified to operate on a PSF--smoothed image.
By smoothing the input image, we reduce the conditions under which the reference image has to be sharpened to match the input image, which is a significant source of false positives.
We note that this technique commutes with the traditional order of operations, in which two images are PSF--matched, then then filtered for detection by the science image PSF.
We compare these two techniques by measuring the empirical rates of false detections in a suite of high--fidelity simulated images from the Large Synoptic Survey Telescope.
We additionally compare the empirical rate to that expected of random fluctuations in the background Gaussian field, including the run of false positives with detection threshold.
In the traditional post--filtering process, deconvolution leads to a significant number of false detections at high signal--to--noise.
In pre--filtering, we find a close match in the number of false detections versus detection threshold, in all conditions.
We estimate that in order for there to be less than 1 expected false detection per LSST sensor, the detection pipeline must operate at a threshold of 5.5--sigma.
This result is seeing dependent.
We found a significant dependence of the false detection rate on our understanding of the image variance; even a 2\% misestimate of the noise may change the false detection rate by a factor of 1.5--2 at 5--sigma.
Pre--filtering is preferred in this regard, as the per--pixel variance is shown to track the empirical variance to higher fidelity at the detection stage.
We find that over or under subtraction of the image background leads to a bias in the positive--going to negative--going false detection ratio, but not on the overall number of detections.
The use of pre--filtering may obviate the need for deconvolution of the template image when the science image has better seeing.
We also present the results of a testing formalism used to tune the configuration of the overall PSF--matching algorithm.
\end{abstract}
\keywords{methods: data analysis, techniques: image processing}

\section{Introduction}

Image subtraction as a solution to general crowded--field variability detection.
PSF-matching as a means to image subtraction.
Sources of false detections (in binary classification these would be naturally referred to as ``false positives''.  However, since ``positive'' also may refer to the polarity of a difference image detection, and we indeed accept both positive--going and negative--going detections as ``false positives'', we choose to use the term ``false detections'' here to refer to these systematic errors).
Machine learning as a way to address false positive rate.
LSST.
Alert stream.
Mention that one paper that talks about pre--filtering but not for this reason.

\section{PSF--Matching Algorithm}

We outline the image subtraction process below.
We assume that science image $S(x,y)$ can be modeled as a convolution of the template image $T(x,y)$ by a PSF--matching kernel $K(u,v;x,y)$ (indices $u,v$ indicate that the kernel itself is a 2--dimensional function, which varies as a function of position $x,y$ in the image.
During convolution and correlation there is an implicit summation over $u,v$).
The images will have different point spread functions (PSFs), which are the time--averaged transfer functions of a point source through the Earth's atmosphere, telescope optics, and into the silicon of the detector before being read out.
The essence of image subtraction is to match the PSFs of these two images so that they may be subtracted pixel by pixel.
We further assume that the PSF--matching kernel may be decomposed using a set of basis functions $K(u,v) = \sum_i a_i K_i(u,v)$, where the coefficients in front of each basis are determined through:
\begin{eqnarray}
\label{eq-soln}
C_i & \equiv & (K_i \otimes T) \\ 
b_{i}  & = & \sum_{x,y} {{C_i(x,y) S(x,y)}\over{\sigma^2(x,y)}}   \nonumber \\
M_{ij} & = & \sum_{x,y} {{C_i(x,y) C_j(x,y)}\over{\sigma^2(x,y)}}  \nonumber \\
a_{i}  & = & M^{-1}_{ij} b_{j} \nonumber
\label{eq-soln}
\end{eqnarray}
Here $\sigma^2(x,y)$ is the per--pixel variance.
To generate a spatially varying model for the kernel, we assume that the relative weights of the basis functions $a_i$ themselves vary spatially, i.e. $K(u,v;x,y) = \sum_i a_i(x,y) K_i(u,v)$.
In general, one may assume that there is a differential background between the two images $b(x,y)$ that may be fit for using a low--order polynomial.
The image difference is then calculated through $D(x,y) = S(x,y) - T(x,y) \otimes K(u,v;x,y) - b(x,y)$.

The basis functions themselves $K_i(u,v)$ are a degree of freedom in this problem.
The original implementations \citep{Alard98,Alard00} used a set of 3 Gaussians, each with a different width, and each modified by a Laguerre polynomial to a given order.
Subsequent studies \citep[e.g.][]{2007AN....328...16I} have suggested that a constant ratio be maintained between the different Gaussian widths, such that $\sigma_{i+1} = \beta \times \sigma_{i}$.  
We use the value $\beta = 2.0$ here.
We set the overall scale for sigma by noting that, under the assumption that the PSFs of the images are Gaussian ($\sigma_S$ for the science image and $\sigma_T$ for the template image), the sigma of the matching kernel should be simply $\sigma_K^2 = \sigma_S^2 - \sigma_T^2$.
We use this width for the central Gaussian in a 3--component basis, with the smaller/larger Gaussian a factor of $1/\beta$ and $\beta$ different, respectively.

Detection on the difference image occurs through correlation of $D(x,y)$ with the science image's PSF, yielding optimally--filtered detection image $D'(x,y) = D(x,y) \circ PSF_S(u,v;x,y)$ where $\circ$ denotes correlation.
The values of the pixels in $D'(x,y)$ provide a maximum likelihood estimate of there being a point source at that position.
Detection occurs by simply finding pixels that are more than N sigma above the square root of the per--pixel variance.

In this work we investigated two orders of operations for the optimal filtering by the PSF.
In the first method, which is the classical post--filtering implementation of the technique, we create a difference image that is then post--filtered with the PSF for detection:
\begin{eqnarray}
D_{Post}(x,y) & = & S(x,y) - T(x,y) \otimes K(u,v;x,y)  \nonumber \\
D'_{Post}(x,y) & = & D_{Post}(x,y) \circ PSF_S(u,v;x,y).  \nonumber
\end{eqnarray}
In the second method, which we refer to as pre--filtering, we PSF--match the template image to a pre--filtered science image:
\begin{eqnarray}
D_{Pre}(x,y) & = & S(x,y) \circ PSF_S(u,v;x,y) - T(x,y) \otimes K(u,v;x,y) \nonumber \\
D'_{Pre}(x,y) & = & D_{Pre}(x,y). \nonumber
\end{eqnarray}
By pre--smoothing the image that we are matching the template to -- $S(x,y) \circ PSF_S(u,v;x,y)$ -- we more frequently avoid the case of PSF--matching {\it deconvolution}, where the full width at half maximum (FWHM) of $T$ is broader than $S$.
This pre--filtering increases $\sigma_S$ by a factor of $\sqrt{2}$.
The PSF matching kernels will need to be correspondingly larger to account for the larger effective PSFs.
We are able to run detection directly on this image, but this requires a modified suite of model--based measurement algorithms.

@@

In order to test the efficacy of these two algorithms, we generated simulated images using the LSST PhoSim package v3.2.5 \cite{phosim}.
These images are high--fidelity simulations of the end--to--end LSST system, including effects from the atmosphere, optics, and detectors.
In order to establish a algorithmic baseline, we turned off certain confounding elements of the simulations, including airglow variations, cloud screens, tracking variations, quantum efficiency variations, diffraction, saturation, and blooming.
An input stellar catalog was generated to seed the simulation that included a random field with of $1000$ stars per sensor, and an equivalent $r$-band magnitude range ($19<r<21$).
All stars in the sample had identical spectral energy distributions, corresponding to a Kurucz--model XXX star \cite{kurucz}, to avoid chromatic effects.
For these sources, proper motion, parallax, and variability were set to zero.
Thus there is no intrinsic variability of any sort designed into the simulated images, and any detections in their difference will be by--definition false positives.

Each simulation run was generated using the following configurations:
\begin{itemize}
\item {\tt Visit 0}: A 300s template image with an atmospheric seeing value of 0.88 arcsec.
\item {\tt Visit 1}: A 15s science image with 0.6 arcsec atmospheric seeing.
\item {\tt Visit 2}: A 15s science image with 0.88 arcsec atmospheric seeing.
\item {\tt Visit 3}: A 15s science image with 1.2 arcsec atmospheric seeing.
\end{itemize}
The seeing of the template image was designed to result in a sharpening deconvolution when compared to {\tt Visit 1}, and a smoothing convolution when compared to {\tt Visit 3}.
All observations were simulated at the zenith, and all images were simulated in the $i$--band.
Sky backgrounds in the resulting images were at the level of $\sim 150$ counts per pixel.
Nine sensors of the central raft were simulated, for a total of 27 difference images per configuration (9 sensors each with 3 seeings to compare to the template image) to use in our comparative analysis.

\section{Optimizing the Algorithm}

To assess the ``quality'' of each implementation of the algorithm, we use as a metric the number of false positive--going and negative--going sources that are detected and measured in each difference image.  We describe our image processing pipeline below, and then describe our assessment and tuning of difference image quality.

% Once we found an optimum configuration, we perturbed the solution along multiple dimensions to understand how the false detection rate responded.

@@

At the start of the pipeline, we have a pair of template and science images that we wish to astrometrically and photometrically register.
The process of astrometric registration refers to the exact alignment of the two images in pixel space, while photometric alignment refers to matching of the PSF shape, sky backgrounds, and zero--point.
From the input reference catalog we choose 80\% of the stars to use in the registration process, with the remaining 20\% of the objects serving as a control sample to assess
the effectiveness of $K(u,v;x,y)$ to subtract objects that were {\it not} used in the fits.
A direct image--to--image relative astrometric mapping was determined using the detected sources in the template and science images, and a sinc--based resampling kernel applied to the template image (as it has larger signal--to--noise than the science image).
The resulting warped image was used as the image subtraction template $T(x,y)$.
The science image $S(x,y)$ (optionally pre-filtered with its $PSF_S(u,v;x,y)$) and $T(x,y)$ were sent to the image subtraction task, whose purpose is to fit for $K(u,v;x,y)$ and to produce the difference image $D(x,y)$.

The first step in this process is to use the PSF Gaussian widths to define the {\it sizes} of the Gaussians' shapes in $K_i(u,v)$.
It is important to note that these widths are {\it not} configurable within the PSF--matching, and must be chosen by the user beforehand.
As mentioned above, the central Gaussian width was chosen to be $\sqrt{\sigma_S^2 - \sigma_T^2}$, and smaller and larger Gaussians used that were scaled a factor of $\beta = 2$.
We modified each of the Gaussians by a set of Laguerre polynomials of a given order.
The smallest Gaussian was modified to the specified order, with the others modified by floor(order/2).
For the process of deconvolution, we followed the prescription from \cite{0266-5611-26-8-085002} to choose basis shapes.
In our implementation of this algorithm, the number of Gaussians is {\it always} fixed to value 3, and their Laguerre degree is always fixed to 3.
The  widths of the Gaussians are determined through the algorithm specified in \cite{0266-5611-26-8-085002}, using as inputs the sequence of Gaussians that we would have used to match a Gaussian of
  width $\sigma_S$ to $\sigma_T$ (i.e. as if we would have convolved the science image and not the template image).
The total number of basis shapes in each set is: $\sum_i^{\rm nGauss} ({\rm degGauss}_i+1)\times({\rm degGauss}_i+2)/2$.

As in the astrometric resampling, we choose to always perform pixel--level operations on the template image given its larger signal--to--noise, and (in the case of real data) fewer artifacts such as dead pixels and cosmic rays.
The dimensions of the PSF matching kernel were chosen to be 6 times the largest Gaussian width, with a minimum kernel size of 21x21 pixels.
For kernels that are significantly smaller than this, the Gaussians have significant (non--zero) power at the kernel boundaries, leading to square systematic artifacts at the scale of the kernel in the difference images.

For each source indexed by $j$, we create substamps from each image $T_j(x,y)$ and $S_j(x,y)$ centered on the source.
These were used to fit for a local solution $K_j(u,v)$; the ensemble of local solutions at each source's $x,y$ was used to fit for the full spatial solution $K(u,v;x,y)$.
We emphasize that the dimensions of the substamps are important.
Note from Equation~\ref{eq-soln} that $C_i$, and thus the coefficients $a_i$, are derived from a convolution of the template substamp with the kernel basis functions $K_i$.
This convolution means that kernel size//2 pixels along the border of each substamp are rendered unusable.
To maintain a significant number of pixels in $C_i$, we set the stamp dimensions to be twice that of the kernel dimensions, such that the number of pixels remaining in $C_i$ is equal to the number of pixels in the kernel.

For each source, we solved for $K_j(u,v)$ to create a local difference image $D_j(x,y)$.
We evaluated several statistics on this local difference image, normalized by the square root of its variance, which puts the pixels in units of $\sigma$.
We measured the mean sigma, the RMS of this distribution, the $\chi^2$ of these pixels normalized by the degrees of freedom (number of pixels in $D_j(x,y)$ minus the number of kernel bases), and the mean squared error.
These are referred to as the {\tt LOCAL} kernel metrics, and reflect how appropriate the chosen kernel basis is.

The ensemble of sources was then used to constrain a spatial model of the kernel $K(u,v;x,y)$.
For the spatial kernel model, we assumed that each of the kernel coefficients $a_i$ may be represented by a $N^{th}$ order 2--dimensional Chebyshev polynomial.
Having generated the full spatial solution $K(u,v;x,y) = \sum_i a_i(x,y) K_i(u,v)$, we evaluated the spatial kernel solution at the position of each source, created a new difference image using this kernel, and recalculated the metrics defined above.
These are referred to as the {\tt SPATIAL} kernel metrics.
Importantly, we also interpolated this solution to the positions of the control sample, which were explicitly {\it not} used in the spatial fit, and evaluated these same metrics.
The differences between the {\tt LOCAL} and {\tt SPATIAL} metrics, as well as the values of the {\tt SPATIAL} metrics for the control sample, reflect the appropriateness of the spatial model and our ability to interpolate and extrapolate to the full extent of the images.
We used the Mean Square Error (MSE) of the control sample to examine the tradeoff between bias and variance in the spatial interpolation.
We define the bias as $\left| data - model \right|$, the variance as $(data - model)^2$, and the MSE as {\tt bias$^2$ + variance}.  
In this context, the bias is the mean of the difference image, and the variance is the mean square of the difference image.
We found that N = 5 or 6 was sufficient to minimize the control sample MSE, or maximize the predictive power of the spatial model without overfitting.

Figure~\ref{fig:1} presents 4 sets of common image subtraction failure modes, and associated poor choices in pipeline configuration that may cause them.
